Supplementary material for "Hack Weeks as a model for Data Science Education and Collaboration"

### 2018-06-21 supplementary_material.tex
More minor corrections

Response rates for NHW (2016: 41 responses; 2017: 45 responses) and GHW (2016: 42 responses; 2017: 41 responses) were 100 percent in both years; the response rate for AHW was 71 percent (35 out of 49) in 2016 and 82 percent (37 out of 45) in 2017. The lower response rates for AHW can be explained by the generally lower rates of response for the time-delayed survey in 2016. In 2017, a number of participants did not attend the full week, and thus several attendees were not present on the last day when the survey was administered. For the AHW surveys, we checked whether the demographics of the respondents diﬀered signiﬁcantly from those of the attendees overall in three relevant categories (career stage, racial/ethnic background, and gender identity) and found no signiﬁcant diﬀerence for any of the groups except for career stage in 2017: Here, we ﬁnd that graduate students are underrepresented (27 percent of respondents compared to 41 percent of attendees), while undergraduate students are overrepresented (25 percent of respondents compared to 12 percent of attendees). Because we group responses in the career stage category into the categories “early-career” (students and postdoctoral fellows), “late-career” (faculty, research scientists), and “other” (e.g. participants from outside academia), this discrepancy has no bearing on the results we report. For gender, we grouped participants into “minority” (women and non-binary) and “non-minority” (men) participants; similarly, for race and ethnicity, we divided participants into the categories “non-white” (all but Caucasian participants) and “white” (Caucasian). Note that demographic data for GHW and NHW is only available for 2016, while for AHW, data is available for both 2016 and 2017. Responding to any question in the survey was optional.

@@

We note that while a non-significant $p$-value can reject a hypothesis, this is not automatically evidence for the corresponding null hypothesis, since a lack of statistical power could also originate in a too small sample size. For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size following~\citep{}. We generate 10,000 random samples via bootstrap for each response variable and compute the corresponding effect size in order to generate 90\% confidence intervals around the measured Cram\'{e}r's V. For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $[0,1]$, we only require an upper bound} this equivalence bound. If the latter hypothesis can be rejected, this indicates that the observed effect falls within the equivalence bounds and is statistically smaller than any effect deemed worthwhile.

@@

Here we present the full results of the statistical analysis performed for the survey (Table \ref{tab:survey}). We tested for correlations between responses to a number of questions and demographic factors, most notably gender identity and racial/ethnic identity. The table also shows (1) the number of responses $N$ to each question for each hack week, (2) value of the $\chi^2$ and Cram\'{e}r's V statistics, $\chi^2$ and $\phi_c$, respectively, (3) the $p$-value associated with the $\chi^2$ statistic, and (4) the degrees of freedom for both $\chi^2$-test ($\mathrm{dof}_{\chi^2}$) and Cram\'{e}r's V ($\mathrm{dof}_{\phi_c}$). The data sets used in this data analysis are presented in Section \ref{sec:suppfigures}, Figures \ref{fig:corr1} through \ref{fig:corr12}

For no combination of questions with a demographic variable, do we ﬁnd a $p$-value below the trial-corrected signiﬁcance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials}. In the main text, we place these null results in context, and refer to the tests that were near significance, and were coupled with a moderate to large effect size, as an indication of where future research may be fruitful. The results also discuss the present study's limitations.

@@

\begin{tabularx}{24cm}{p{3.2cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}}

@@

\textbf{Question} & \multicolumn{9}{c}{\textbf{Astro Hack Week}} & \multicolumn{9}{c}{\textbf{GeoHackWeek}} & \multicolumn{9}{c}{\textbf{NeuroHackWeek}} \\ \midrule
 & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
 & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ &$\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
 & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ &$p_{\mathrm{eq}}$  \\ \midrule

@@

New topics, tools, methods (a) & 	61 & 10 &   5.64 & 0.85 & 1.93 & 0      & 0 & 0.26 & 0.12 &	36 & 6   &   2.51 & 0.87 & 1.88 & 0      & 0 & 0.31 & 0.16 &		41 & 10 & 15.53 & 0.15 & 1.90 & 0.23 & 0 & 0.32 & 0.17 \\
Improvement of research (b) & 		61 &   6 &   9.89 & 0.13 & 1.93 & 0.18 & 0 & 0.24 & 0.08 &	35 & 4   & 12.41 & 0.02 & 1.88 & 0.35 & 0 & 0.28 & 0.13 &		40 &   6 &   7.42 & 0.28 & 1.90 & 0.13 & 0 & 0.31 & 0.14 \\
Better science (c) & 				59 &   6 &   3.79 & 0.70 & 1.93 & 0      & 0 & 0.25 & 0.09 &	35 & 4   & 10.61 & 0.03 & 1.88 & 0.31 & 0 & 0.27 & 0.12 & 		40 &   4 &   5.06 & 0.29 & 1.90 & 0.11 & 0 & 0.28 & 0.12 \\
Teaching (d) & 					61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 0 & 0.28 & 0.13 &	36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 0 & 0.34 & 0.20 & 		40 & 10 &   6.90 & 0.74 & 1.90 &      0 & 0 & 0.33 & 0.19 \\ \midrule

@@

Improvement of research (b) & 		60 &   3 & 2.08 & 0.56 & 0.98 & 0      & 0 & 0.26 & 0.02&		35 & 2 &   5.26 & 0.07 & 0.97 & 0.30 & 0      & 0.33 & 0.07 &		40 & 3 & 7.62 & 0.05 & 0.97 & 0.34 & 0 & 0.30 & 0.05\\
Teaching (d) & 					61 &   5 & 7.30 & 0.20 & 0.98 & 0.19 & 0 & 0.30 & 0.05 & 		36 & 5 & 12.71 & 0.26 & 0.97 & 0      & 0.37 & 0.46 & 0.11 & 		40 & 5 & 3.17 & 0.67 & 0.97 & 0      & 0 & 0.36 & 0.11\\
Connection building (e) & 			61 &   4 & 3.70 & 0.45 & 0.98 & 0      & 0 & 0.28 & 0.03 & 		36 & 3 &   5.89 & 0.12 & 0.97 & 0.28 & 0      & 0.34 & 0.08 &		41 & 3 & 4.42 & 0.22 & 0.98 & 0.18 & 0 & 0.32 & 0.07 \\
Valued contributions (f) & 			57 &   4 & 5.98 & 0.20 & 0.98 & 0.18 & 0 & 0.28 & 0.03 &		34 & 5 &   9.46 & 0.09 & 0.97 & 0.36 & 0      & 0.34 & 0.08 &		38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 & 0 & 0.35 & 0.10\\ \midrule

@@

Improvement of research (b)  & 	59 &   3 & 0.97   & 0.81 & 0.98 & 0      & 0 & 0.27 & 0.03 &	34 & 2 & 0.01 & 0.99 & 0.97 & 0      & 0 & 0.34 & 0.08 &			39 & 3 & 2.79 & 0.43 & 0.97 & 0      & 0 & 0.32 & 0.06 \\
Teaching (d) & 					61 &   5 & 5.63   & 0.34 & 0.98 & 0.09 & 0 & 0.31 & 0.05 &	34 & 4 & 1.29 & 0.86 & 0.97 & 0      & 0 & 0.39 & 0.12 & 			39 & 5 & 3.06 & 0.69 & 0.97 & 0      & 0 & 0.38 & 0.12 \\
Connection building (e) & 			61 &   4 & 1.64   & 0.80 & 0.98 & 0      & 0 & 0.28 & 0.04 &	34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 0 & 0.36 & 0.09 &			40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 & 0 & 0.33 & 0.07 \\
Valued contributions (f) & 			58 &   4 & 13.14 & 0.01 & 0.98 & 0.40 & 0 & 0.29 & 0.04 &	32 & 5 & 2.49 & 0.78 & 0.96 & 0      & 0 & 0.40 & 0.14 &			37 & 5 & 4.05 & 0.54 & 0.97 & 0      & 0 & 0.36 & 0.11\\

@@

The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear.
Individuals with IS thus experience a fear of being ``found out'', shamed and expelled from their environment \citep{Clance1978-ef}.
Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS, namely that participants feel the need to be experts in multiple scientific and technical aspects of the activities pursued during a hack week.
This particular form of IS hinges to some degree on the design focus on diversity of backgrounds in that all the other participants seem to know something that any given participant does not. This might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee.
This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \citep{Parkman2016-ro}.
Less severely, another major concern is that IS inhibits risk taking: Participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, and to be pro-active about forming new collaborations.
 Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, and community building are hampered by IS.
 Thus, we are working within our hack weeks to mitigate IS using various techniques.
With respect to minority participants, ensuring adequate representation can decrease feelings of otherness and may help reduce IS.
More generally, being open about the presence and prevalence of IS can help participants feel more at ease. %(e.g. one Astro Hack Week participant remarked in a survey, \textit{''I really appreciated the direct acknowledgement of impostor syndrome on the first day.
%I think it helped ease the feeling!!``}).
Additionally, role models are very effective at encouraging positive behavior: Asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking. We also invited our organizing committees to model open communication about their own handling of IS through the course of their career. We found that transparency about one's own uncertainty, especially from community leaders, helped create a hospitable atmosphere and removed barriers.

>>>>

Response rates for NHW (2016: 41 responses; 2017: 45 responses) and GHW (2016: 42 responses; 2017: 41 responses) were 100 percent in both years; the response rate for AHW was 71 percent (35 out of 49) in 2016 and 82 percent (37 out of 45) in 2017. The lower response rates for AHW can be explained by the generally lower rates of response for the time-delayed survey in 2016. In 2017, a number of participants did not attend the full week, and thus several attendees were not present on the last day when the survey was administered. For the AHW surveys, we checked whether the demographics of the respondents diffÄered signifÅcantly from those of the attendees overall in three relevant categories (career stage, racial/ethnic background, and gender identity) and found no signifiÅcant difference for any of the groups except for career stage in 2017: Here, we find that graduate students are underrepresented (27 percent of respondents compared to 41 percent of attendees), while undergraduate students are overrepresented (25 percent of respondents compared to 12 percent of attendees). Because we group responses in the career stage category into the categories ``Äúearly-career''Äù (students and postdoctoral fellows), ``late-career'' (faculty, research scientists), and ``other"Äù (e.g. participants from outside academia), this discrepancy has no bearing on the results we report. For gender, we grouped participants into ``Äúminority'' (women and non-binary) and ``non-minority''Ä (men) participants; similarly, for race and ethnicity, we divided participants into the categories ``non-white'' (all but Caucasian participants) and ``white'' (Caucasian). Note that demographic data for GHW and NHW is only available for 2016, while for AHW, data is available for both 2016 and 2017. Responding to any question in the survey was optional.

@@

We note that while a non-significant $p$-value can reject a hypothesis, this is not automatically evidence for the corresponding null hypothesis, since a lack of statistical power could also originate in a too small sample size. For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size following~\citep{shiskina2018}. We generate 10,000 random data sets from the measured responses by sampling with replacement for each response variable and compute the corresponding effect size in order to generate confidence intervals around the measured Cram\'{e}r's V. For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $[0,1]$, we only require an upper bound} this equivalence bound. If the latter hypothesis can be rejected, this indicates that the observed effect falls within the equivalence bounds and is statistically smaller than any effect deemed worthwhile.

@@

Here we present the full results of the statistical analysis performed for the survey (Table \ref{tab:survey}). We tested for correlations between responses to a number of questions and demographic factors, most notably gender identity and racial/ethnic identity. The table also shows (1) the number of responses $N$ to each question for each hack week, (2) value of the $\chi^2$ and Cram\'{e}r's V statistics, $\chi^2$ and $\phi_c$, respectively, (3) the $p$-value associated with the $\chi^2$ statistic, (4) the degrees of freedom for both $\chi^2$-test ($\mathrm{dof}_{\chi^2}$) and Cram\'{e}r's V ($\mathrm{dof}_{\phi_c}$), and (5) the upper 95\% confidence interval for the Cram\'{e}r's V statistic ($\mathrm{CI}_{\mathrm{U}}$) as well as the corresponding $p$-value that the null hypothesis (the presence of an effect with moderate or higher strength) can be rejected, $p_\mathrm{eq}$ . The data sets used in this data analysis are presented in Section \ref{sec:suppfigures}, Figures \ref{fig:corr1} through \ref{fig:corr12}

For no combination of questions with a demographic variable, do we find a $p$-value below the trial-corrected significance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials}. In the main text, we place these null results in context, and refer to the tests that were near significance, and were coupled with a moderate to large effect size, as an indication of where future research may be fruitful. The results also discuss the present study's limitations.

@@

\begin{tabularx}{24cm}{p{4.1cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{1.0cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{1.0cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{1.0cm}}

@@

\textbf{Question} & \multicolumn{8}{c}{\textbf{Astro Hack Week}} & \multicolumn{8}{c}{\textbf{GeoHackWeek}} & \multicolumn{8}{c}{\textbf{NeuroHackWeek}} \\ \midrule
 & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
 & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
 & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{U}}$ &$p_{\mathrm{eq}}$  \\ \midrule

@@

New topics, tools, methods (a) & 	61 & 10 &   5.64 & 0.85 & 1.93 & 0      & 0.26 & 0.12 &	36 & 6   &   2.51 & 0.87 & 1.88 & 0      & 0.31 & 0.16 &		41 & 10 & 15.53 & 0.15 & 1.90 & 0.23 & 0.32 & 0.17 \\
Improvement of research (b) & 		61 &   6 &   9.89 & 0.13 & 1.93 & 0.18 & 0.24 & 0.08 &	35 & 4   & 12.41 & 0.02 & 1.88 & 0.35 & 0.28 & 0.13 &		40 &   6 &   7.42 & 0.28 & 1.90 & 0.13 & 0.31 & 0.14 \\
Better science (c) & 				59 &   6 &   3.79 & 0.70 & 1.93 & 0      & 0.25 & 0.09 &	35 & 4   & 10.61 & 0.03 & 1.88 & 0.31 & 0.27 & 0.12 & 		40 &   4 &   5.06 & 0.29 & 1.90 & 0.11 & 0.28 & 0.12 \\
Teaching (d) & 					61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 0.28 & 0.13 &	36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 0.34 & 0.20 & 		40 & 10 &   6.90 & 0.74 & 1.90 &      0 & 0.33 & 0.19 \\ \midrule

@@

Improvement of research (b) & 		60 &   3 & 2.08 & 0.56 & 0.98 & 0      & 0.26 & 0.02&		35 & 2 &   5.26 & 0.07 & 0.97 & 0.30 & 0.33 & 0.07 &		40 & 3 & 7.62 & 0.05 & 0.97 & 0.34 & 0.30 & 0.05\\
Teaching (d) & 					61 &   5 & 7.30 & 0.20 & 0.98 & 0.19 & 0.30 & 0.05 & 	36 & 5 & 12.71 & 0.26 & 0.97 & 0      & 0.46 & 0.11 & 		40 & 5 & 3.17 & 0.67 & 0.97 & 0      & 0.36 & 0.11\\
Connection building (e) & 			61 &   4 & 3.70 & 0.45 & 0.98 & 0      & 0.28 & 0.03 & 	36 & 3 &   5.89 & 0.12 & 0.97 & 0.28 & 0.34 & 0.08 &		41 & 3 & 4.42 & 0.22 & 0.98 & 0.18 & 0.32 & 0.07 \\
Valued contributions (f) & 			57 &   4 & 5.98 & 0.20 & 0.98 & 0.18 & 0.28 & 0.03 &	34 & 5 &   9.46 & 0.09 & 0.97 & 0.36 & 0.34 & 0.08 &		38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 & 0.35 & 0.10\\ \midrule

@@

Improvement of research (b)  & 	59 &   3 & 0.97   & 0.81 & 0.98 & 0      & 0.27 & 0.03 &	34 & 2 & 0.01 & 0.99 & 0.97 & 0      & 0.34 & 0.08 &			39 & 3 & 2.79 & 0.43 & 0.97 & 0      & 0.32 & 0.06 \\
Teaching (d) & 					61 &   5 & 5.63   & 0.34 & 0.98 & 0.09 & 0.31 & 0.05 &	34 & 4 & 1.29 & 0.86 & 0.97 & 0      & 0.39 & 0.12 & 			39 & 5 & 3.06 & 0.69 & 0.97 & 0      & 0.38 & 0.12 \\
Connection building (e) & 			61 &   4 & 1.64   & 0.80 & 0.98 & 0      & 0.28 & 0.04 &	34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 0.36 & 0.09 &			40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 & 0.33 & 0.07 \\
Valued contributions (f) & 			58 &   4 & 13.14 & 0.01 & 0.98 & 0.40 & 0.29 & 0.04 &	32 & 5 & 2.49 & 0.78 & 0.96 & 0      & 0.40 & 0.14 &			37 & 5 & 4.05 & 0.54 & 0.97 & 0      & 0.36 & 0.11\\

@@

The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear.
Individuals with IS thus experience a fear of being ``found out'', shamed and expelled from their environment \citep{Clance1978-ef}.
Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS, namely that participants feel the need to be experts in multiple scientific and technical aspects of the activities pursued during a hack week.
This particular form of IS hinges to some degree on the design focus on diversity of backgrounds in that all the other participants seem to know something that any given participant does not. This might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee.
This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \citep{Parkman2016-ro}.
Less severely, another major concern is that IS inhibits risk taking: Participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, and to be pro-active about forming new collaborations.
 Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, and community building are hampered by IS.
 Thus, we are working within our hack weeks to mitigate IS using various techniques.
With respect to minority participants, ensuring adequate representation can decrease feelings of otherness and may help reduce IS.
More generally, being open about the presence and prevalence of IS can help participants feel more at ease. Additionally, role models are very effective at encouraging positive behavior: Asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking. We also invited our organizing committees to model open communication about their own handling of IS through the course of their career. We found that transparency about one's own uncertainty, especially from community leaders, helped create a hospitable atmosphere and removed barriers.
